---
title: "Final Project"
author: "Soo-Min Kim, Ibrahim Osman, Tien Tang"
date: today
format:
  html:
    embed-resources: true
---

# Section I:

## Loading Libraries

```{r}
library(dplyr)
library(ggplot2)
```

## Dataset 1:

### Loading Dataset:

```{r}
bigfoot_df <- read.csv("bfro_all_states_county_totals.csv")

glimpse(bigfoot_df)
```

### About the Data:

Bigfoot Sightings Data Source: Bigfoot Field Researchers Organization (BFRO) Website: <https://www.bfro.net/GDB/> State listings: <https://www.bfro.net/GDB/state_listing.asp> Individual report pages describe: Report number Date of sighting County State Season Classification (Class A/B/C) Summary narrative Investigator comments There is no formal downloadable codebook provided by BFRO. Column definitions are inferred from webpage structure. Because the dataset is scraped from HTML pages, variable descriptions must be manually interpreted. The most relevant data for our Bigfoot data is State County Year report_id.

# Section II:

## Dataset 2:

### Loading Dataset:

```{r}
census_df <- read.csv("co-est2023-alldata.csv")

glimpse(census_df)
```

### About the Data: 

Population Data If you used U.S. Census data: Source: United States Census Bureau Data portal: <https://data.census.gov/> County population estimates documentation: <https://www.census.gov/programs-surveys/popest.html> The Census Bureau provides formal metadata and documentation for: County name FIPS code Year Total population Demographic breakdowns The relevant data set for the population County State Year Total population.

# Section III:

## Joining the Data

```{r}
state_lookup <- data.frame(
  state_name = state.name,
  state_abbr = state.abb
)


bigfoot_df <- bigfoot_df |>
  left_join(state_lookup, by = c("state" = "state_abbr"))

#Make a common column to connect 
census_df <- census_df |>
  mutate(state_name = STNAME)


census_state <- census_df |>
  group_by(state_name) |>
  summarize(
    population = sum(POPESTIMATE2023, na.rm = TRUE)
  )

join_df <- left_join(bigfoot_df, census_state, by = "state_name")

glimpse(join_df)
```

## Data Documentation:

Our combined dataset includes reported Bigfoot sightings from the BFRO and county-level population data from the U.S. Census Bureau. A major limitation is that Bigfoot sightings are self-reported and voluntary, meaning the data likely reflects reporting bias rather than true geographic distribution. Counties with higher internet access, stronger folklore culture, or more active local investigators may appear to have more sightings. Additionally, the dataset does not include demographic information about reporters, which limits analysis of social or cultural influences. The U.S. Census Bureau also doesnt provide an accurate list of population per county because it’s mostly based on estimates. In relation to the eight human rights principles, the dataset partially reflects transparency and accountability because Census data is well-documented, but BFRO data lacks formal documentation. Privacy risks are low since no personal identifiers are used. However, fairness and non-discrimination may be affected if rural or low-access communities are underrepresented. Professional responsibility requires acknowledging these biases and clearly communicating uncertainty when presenting results.

# Section IV:

## Data Wrangle 1:

### Cleaning Data Through Removing NAs:

```{r}
join_df <- join_df |>
  filter(!is.na(listings))

glimpse(join_df)

```

## Data Wrangle 2:

### Cleaning Data Through Removal of Unnecessary Columns:

```{r}
join_df <- join_df |>
  select(-Last.Posted, -state_name)

glimpse(join_df)
```

## Data Wrangle 3:

### Graphing Data:

```{r}
ggplot(join_df, aes(x = population, y = listings)) +
  geom_point() +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    x = "State Population (2023)",
    y = "Bigfoot Sightings",
    title = "Bigfoot Sightings vs State Population"
  )
```

# Section V:

## Group Information TA: Section BF Marco Guo

## Wed 11:30-12:20 PM 

## Group Overview:

Our group’s main topic examines the relationship between Bigfoot sightings and population size across the United States, we did this by reporting each sighting as a data point in order to help us report trends rather than proof of Bigfoot’s existence. By integrating sighting data with demographic statistics, we aim to determine whether there is a correlation between the number of reported sightings and population density. In specific, we investigated whether or not sightings were more prevalent in denser populations, primarily in urban and rural areas in which the legend is traditionally associated with. We did this by shifting the focus from the creature itself to the people reporting it.

## Group Members

### Ibrahim Osman: 

Affiliation: Pre-major (Undeclared), University of Washington : iosman\@uw.edu

Ibrahim is interested primarily in analyzing just how the demographic patterns actually influence the reported Bigfoot sightings, Ibrahim is interested in what these patterns reveal about the creature as well as about human perception. He is interested informatics simply because he wants to better understand how data analysis can uncover trends in social behavior.

### Soo-Min Kim: 

Affiliation: Real Estate, Universtiy of Washington : skim0326\@uw.edu

Soo-Min is interested in how population density could affect the frequency of the sightings. In Informatics, he is interested in the organizational aspect of informatics and how data can affect people and things that rely on data.

### Tien Tang: 

Affiliation: Pre-major (Undeclared), University of Washington : Tangco15\@uw.edu

Tien Tang is mainly focused on examining how population demographics shape the patterns of reported Bigfoot sightings. He wants to explore what these trends suggest not only about the legend itself, but also about the ways people interpret and report unusual experiences. His interest in informatics comes from a desire to use data driven methods to uncover meaningful insights about human behavior and collective perception.

## Coding Notes: 

For this group project, we used the tools taught to such in class such as Jupyter Hub and RStudio, to help with collecting the Data, organizing the data as well as simplifying the data in an orderly manner. We primary relied on class lectures and our individual lab exercises in order to complete our coding tasks. There were other websites like stackoverflow.com to supplement areas that we were lacking such as the specific number change on the graph or the setup of the datasets for them to be able to merge.
